A density ratio framework for evaluating the utility of synthetic data

### 2024-10-15 paper/paper.qmd
update paper
small change abstract + cover letter

---
title: "A density ratio framework for evaluating the utility of synthetic data"
format:
  arxiv-pdf:
    keep-tex: true
    include-in-header:
      - file: preamble.tex
    header-includes:
      - \usepackage{float}
      - \floatplacement{table}{t}
  arxiv-html: default
number-sections: true
linenumbers: false
doublespacing: true
runninghead: "Density ratios for utility"
authorcols: true
link-citations: true
author:
  - name: Thom Benjamin Volker
    affiliations:
      - name: Utrecht University | Statistics Netherlands
        department: Methodology and Statistics | Methodology
        address: Padualaan 14
        city: Utrecht
        country: The Netherlands
        postal-code: "3584CH"
    orcid: 0000-0002-2408-7820
    email: t.b.volker@uu.nl
  - name: Peter-Paul de Wolf
    affiliations:
      - name: Statistics Netherlands
        department: Methodology
        address: PO Box 24500
        city: The Hague
        country: The Netherlands
        postal-code: "2490HA"
    email: pp.dewolf@cbs.nl
  - name: Erik-Jan van Kesteren
    affiliations:
      - name: Utrecht University
        department: Methodology and Statistics
        address: Padualaan 14
        city: Utrecht
        country: The Netherlands
        postal-code: "3584CH"
    email: e.vankesteren1@uu.nl
abstract: | 
  Synthetic data generation is a promising technique to facilitate the use of sensitive data while mitigating the risk of privacy breaches. However, for synthetic data to be useful in downstream analysis tasks, it needs to be of sufficient quality. Various methods have been proposed to measure the utility of synthetic data, but their results are often incomplete or even misleading. In this paper, we propose using density ratio estimation to improve quality evaluation for synthetic data, and thereby the quality of synthesized datasets.
  We show how this framework relates to and builds on existing measures, yielding global and local utility measures that are informative and easy to interpret. We develop an estimator which requires little to no manual tuning due to automatic selection of a nonparametric density ratio model. Through simulations, we find that density ratio estimation yields more accurate estimates of global utility than established procedures. A real-world data application demonstrates how the density ratio can guide refinements of synthesis models and can be used to improve downstream analyses. We conclude that density ratio estimation is a valuable tool in synthetic data generation workflows and provide these methods in the accessible open source R-package densityratio.
keywords: |
  synthetic data, utility, density ratio, privacy, disclosure limitation
bibliography: bibliography.bib
params:
  runsims: false
---

>>>>

---
title: "A density ratio framework for evaluating the utility of synthetic data"
format:
  arxiv-pdf:
    keep-tex: true
    include-in-header:
      - file: preamble.tex
    header-includes:
      - \usepackage{float}
      - \floatplacement{table}{t}
  arxiv-html: default
number-sections: true
linenumbers: false
doublespacing: true
runninghead: "Density ratios for utility"
authorcols: true
link-citations: true
author:
  - name: Thom Benjamin Volker
    affiliations:
      - name: Utrecht University | Statistics Netherlands
        department: Methodology and Statistics | Methodology
        address: Padualaan 14
        city: Utrecht
        country: The Netherlands
        postal-code: "3584CH"
    orcid: 0000-0002-2408-7820
    email: t.b.volker@uu.nl
  - name: Peter-Paul de Wolf
    affiliations:
      - name: Statistics Netherlands
        department: Methodology
        address: PO Box 24500
        city: The Hague
        country: The Netherlands
        postal-code: "2490HA"
    email: pp.dewolf@cbs.nl
  - name: Erik-Jan van Kesteren
    affiliations:
      - name: Utrecht University
        department: Methodology and Statistics
        address: Padualaan 14
        city: Utrecht
        country: The Netherlands
        postal-code: "3584CH"
    email: e.vankesteren1@uu.nl
abstract: | 
  Synthetic data generation is a promising technique to facilitate the use of sensitive data while mitigating the risk of privacy breaches. However, for synthetic data to be useful in downstream analysis tasks, it needs to be of sufficient quality. Various methods have been proposed to measure the utility of synthetic data, but their results are often incomplete or even misleading. In this paper, we propose to use the density ratio estimation framework to improve quality evaluation for synthetic data, and thereby improve the quality of synthesized datasets. We show how this framework relates to and builds on existing measures, yielding global and local utility measures that are informative and easy to interpret. The proposed methodology employs automatic, non-parametric model selection for the density ratio, and thus requires little to no manual model specification. Through simulations, we find that density ratio estimation yields more accurate estimates of global utility than established procedures. A real-world data application demonstrates how the density ratio can guide refinements of synthesis models and can be used to improve downstream analyses. We conclude that density ratio estimation is a valuable tool in synthetic data generation workflows and make the methodology available through the open source R-package densityratio.
keywords: synthetic data, utility, density ratio, privacy, disclosure limitation
bibliography: bibliography.bib
params:
  runsims: false
---

### 2024-10-15 paper/paper.tex
update paper
small change abstract + cover letter

\usepackage{setspace}
\doublespacing
\newcommand{\runninghead}{A Preprint }
\renewcommand{\runninghead}{Density ratios for utility }
\title{A density ratio framework for evaluating the utility of synthetic
data}
\def\asep{\\\\\\ } % default: all authors on same column
\def\asep{\And }
\author{\textbf{Thom Benjamin
Volker}~\orcidlink{0000-0002-2408-7820}\\Methodology and Statistics
\textbar{} Methodology\\Utrecht University \textbar{} Statistics
Netherlands\\Utrecht,\ 3584CH\\\href{mailto:t.b.volker@uu.nl}{t.b.volker@uu.nl}\asep\textbf{Peter-Paul
de Wolf}\\Methodology\\Statistics Netherlands\\The
Hague,\ 2490HA\\\href{mailto:pp.dewolf@cbs.nl}{pp.dewolf@cbs.nl}\asep\textbf{Erik-Jan
van Kesteren}\\Methodology and Statistics\\Utrecht
University\\Utrecht,\ 3584CH\\\href{mailto:e.vankesteren1@uu.nl}{e.vankesteren1@uu.nl}}
\date{}
\begin{document}
\maketitle
\begin{abstract}
Synthetic data generation is a promising technique to facilitate the use
of sensitive data while mitigating the risk of privacy breaches.
However, for synthetic data to be useful in downstream analysis tasks,
it needs to be of sufficient quality. Various methods have been proposed
to measure the utility of synthetic data, but their results are often
incomplete or even misleading. In this paper, we propose using density
ratio estimation to improve quality evaluation for synthetic data, and
thereby the quality of synthesized datasets. We show how this framework
relates to and builds on existing measures, yielding global and local
utility measures that are informative and easy to interpret. We develop
an estimator which requires little to no manual tuning due to automatic
selection of a nonparametric density ratio model. Through simulations,
we find that density ratio estimation yields more accurate estimates of
global utility than established procedures. A real-world data
application demonstrates how the density ratio can guide refinements of
synthesis models and can be used to improve downstream analyses. We
conclude that density ratio estimation is a valuable tool in synthetic
data generation workflows and provide these methods in the accessible
open source R-package densityratio.
\end{abstract}
{\bfseries \emph Keywords}
\def\sep{\textbullet\ }
synthetic data, utility, density ratio, privacy, disclosure
limitation \sep 
synthetic data, utility, density ratio, privacy, disclosure limitation

@@

\caption{\label{tbl-application-regression}TODO}

>>>>

\usepackage{setspace}
\doublespacing
\newcommand{\runninghead}{A Preprint }
\renewcommand{\runninghead}{Density ratios for utility }
\title{A density ratio framework for evaluating the utility of synthetic
data}
\def\asep{\\\\\\ } % default: all authors on same column
\def\asep{\And }
\author{\textbf{Thom Benjamin
Volker}~\orcidlink{0000-0002-2408-7820}\\Methodology and Statistics
\textbar{} Methodology\\Utrecht University \textbar{} Statistics
Netherlands\\Utrecht,\ 3584CH\\\href{mailto:t.b.volker@uu.nl}{t.b.volker@uu.nl}\asep\textbf{Peter-Paul
de Wolf}\\Methodology\\Statistics Netherlands\\The
Hague,\ 2490HA\\\href{mailto:pp.dewolf@cbs.nl}{pp.dewolf@cbs.nl}\asep\textbf{Erik-Jan
van Kesteren}\\Methodology and Statistics\\Utrecht
University\\Utrecht,\ 3584CH\\\href{mailto:e.vankesteren1@uu.nl}{e.vankesteren1@uu.nl}}
\date{}
\begin{document}
\maketitle
\begin{abstract}
Synthetic data generation is a promising technique to facilitate the use
of sensitive data while mitigating the risk of privacy breaches.
However, for synthetic data to be useful in downstream analysis tasks,
it needs to be of sufficient quality. Various methods have been proposed
to measure the utility of synthetic data, but their results are often
incomplete or even misleading. In this paper, we propose to use the
density ratio estimation framework to improve quality evaluation for
synthetic data, and thereby improve the quality of synthesized datasets.
We show how this framework relates to and builds on existing measures,
yielding global and local utility measures that are informative and easy
to interpret. The proposed methodology employs automatic, non-parametric
model selection for the density ratio, and thus requires little to no
manual model specification. Through simulations, we find that density
ratio estimation yields more accurate estimates of global utility than
established procedures. A real-world data application demonstrates how
the density ratio can guide refinements of synthesis models and can be
used to improve downstream analyses. We conclude that density ratio
estimation is a valuable tool in synthetic data generation workflows and
make the methodology available through the open source R-package
densityratio.
\end{abstract}
{\bfseries \emph Keywords}
\def\sep{\textbullet\ }
synthetic data, utility, density ratio, privacy, disclosure
limitation \sep 
synthetic data, utility, density ratio, privacy, disclosure limitation

@@

\caption{\label{tbl-application-regression}Linear regression
coefficients for the observed data, synthetic data and reweighted
synthetic data using the density ratio as importance weights. Boldface
text indicates which coefficient has smaller absolute bias relative to
the observed data.}
